---
title: "Two-parameter ridge estimation of HAR realized-volatility models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Two-parameter ridge estimation of HAR realized-volatility models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7,
                      fig.height = 4.5)
library(tpridge)
```

## What the package does

Ridge regression trades a little bias for a large reduction in variance,
which is why it helps when regressors are collinear. The price is that
it fits the estimation sample worse than least squares. The two-parameter
estimator of Lipovetsky and Conklin (2005) recovers part of that lost fit
by rescaling the ridge coefficient vector by a scalar $q$,

$$\hat\beta_{II}(\lambda) = q(\lambda)\,\hat\beta_{I}(\lambda),
\qquad q = \frac{\hat\beta_I' r}{\hat\beta_I' G \hat\beta_I},$$

where $G = X'X$ and $r = X'y$ on the standardized, centred design. This
package implements the estimator, the four analytical results that
characterise it, and the full pipeline from downloading intraday prices
to evaluating forecasts.

Four properties are worth stating up front, because the package checks
all of them and they are what the estimator is for.

**The scalar always exceeds one.** For any $\lambda>0$,
$q = 1 + \lambda\|\hat\beta_I\|^2 / \|X\hat\beta_I\|^2 > 1$. This is
derived rather than assumed.

**The fit gain has a closed form.** The improvement in the coefficient of
determination is exactly
$(q-1)^2\|X\hat\beta_I\|^2 / y'y$, so it can be computed without
re-estimating anything.

**No estimator beats least squares in sample.** The ordering
$R^2(\hat\beta_{OLS}) \ge R^2(\hat\beta_{II}) \ge R^2(\hat\beta_{I})$
holds always. The two-parameter estimator closes part of a gap; it never
crosses it.

**The two estimators fail differently.** As $\lambda\to\infty$ the
one-parameter estimator collapses to the null predictor, while the
two-parameter estimator converges to a finite non-zero limit. Under a
badly chosen penalty this difference is large.

## The estimators on any design

The three core functions take a matrix and a vector, so they work on any
regression problem.

```{r}
set.seed(1)
X <- matrix(rnorm(400), ncol = 4)
y <- as.numeric(X %*% c(1, 0.6, 0.3, 0.1)) + rnorm(100)

tpr_lambda(X, y)                                # data-driven penalty
tpr_q(X, y, lambda = 20)                        # the scalar, standalone
tpr_lambda(X, y, rule = "df", target_df = 2)    # penalty for a target df
```

The penalty is not scale free: the eigenvalues of $X'X$ grow with the
sample, so the same $\lambda$ means different amounts of shrinkage in
different designs. The effective degrees of freedom,
$\mathrm{df}(\lambda)=\sum_j e_j/(e_j+\lambda)$, are comparable, which is
why `rule = "df"` exists.

```{r}
c(tpr_eff_df(X, 0), tpr_eff_df(X, 20), tpr_eff_df(X, 1e6))
```

Fitting returns all three estimators from one decomposition.

```{r}
fit <- tpr_fit(X, y, lambda = 20)
fit
```

The Proposition 2 residual printed above is the empirical fit gain minus
its closed form. It is zero to machine precision, always.

```{r}
plot(fit, which = "path")
```

## From Binance to realized variance

Downloading is a single call. The monthly archives are public and no key
is needed; they are cached, so a month already on disk is not fetched
again.

```{r eval = FALSE}
k <- binance_klines("BTCUSDT", from = "2024-01", to = "2024-06")
rm_ <- realized_measures(k)
head(rm_[, c("date", "n_5m", "RV", "BV", "Jump")])
```

`realized_measures()` returns realized variance, bipower variation, the
jump component, the two semivariances, realized quarticity, skewness,
kurtosis, the Parkinson range and the activity variables. Two
conventions matter. Returns are differenced over the full series rather
than within each calendar day, so the interval spanning midnight is
retained: these markets trade continuously and excluding it would discard
one return per day. And days with fewer than ninety per cent of the
expected intervals have their measures set to `NA` rather than being
silently computed from partial data.

The identities hold by construction, which is a useful check on any
replication:

```{r eval = FALSE}
max(abs(rm_$RV - (rm_$RSV_pos + rm_$RSV_neg)), na.rm = TRUE)   # ~1e-16
```

## Building the HAR model

```{r}
rv <- har_simulate(800, sigma2 = 0.5)
feat <- har_features(rv)          # daily, weekly, monthly, and the lead
des  <- har_design(feat)
dim(des$X)
```

The weekly and monthly components are rolling averages ending at $t-1$,
so every row pairs a forecast origin with the value it predicts. The
logarithmic transformation is the default because a squared error
criterion applied to realized variance in levels is dominated by a
handful of days.

Additional predictors are appended through `extra`:

```{r eval = FALSE}
feat <- har_features(rm_$RV, date = rm_$date,
                     extra = rm_[, c("BV", "Jump", "RSV_neg")])
```

## Rolling forecasts

```{r}
roll <- tpr_rolling(des$X, des$y, window = 365)
summary(roll)
```

Supplying several penalty targets applies all of them to the same window,
so the comparison across penalties is paired and not contaminated by
sampling variation in the data.

```{r}
path <- tpr_rolling(des$X, des$y, window = 365, rule = "df",
                    target_df = c(3, 2.5, 2, 1.5, 1, 0.5))
plot(path, which = "path")
```

The shape of that plot is the practical argument for the second
parameter. The one-parameter estimator reaches an interior optimum and
then deteriorates, in places below least squares; the two-parameter
estimator is close to monotone. Since the appropriate penalty is not
observable in advance, an estimator that degrades gracefully is worth
more than one that attains a marginally higher optimum and collapses away
from it.

```{r}
plot(roll, which = "q")
```

## Evaluation

```{r}
w <- data.table::dcast(roll, origin + actual + prev_mean ~ estimator,
                       value.var = "pred")
tpr_metrics(w$actual, w$`Ridge-II`, benchmark = w$prev_mean)
tpr_dm((w$actual - w$OLS)^2, (w$actual - w$`Ridge-II`)^2)
```

A positive Diebold and Mariano statistic means the first loss series is
larger, so the second model forecasts better.

For tail risk, `tpr_var_backtest()` builds
$\widehat{VaR}_{t+1}=z_\alpha\sqrt{\hat f_{t+1}}$ and applies the
unconditional coverage test of Kupiec and the independence test of
Christoffersen. On cryptocurrency data the Gaussian quantile is rejected
at the one per cent level while a standardised Student-$t$ quantile is
not, which is a distributional failure rather than a failure of the
variance forecast.

```{r eval = FALSE}
tpr_var_backtest(variance_forecast, returns, alpha = 0.01, dist = "student")
```

The model confidence set is available through `tpr_mcs()` when the `MCS`
package is installed.

## Reproducing the Monte Carlo study

```{r eval = FALSE}
mc <- tpr_montecarlo(n = c(100, 200, 400, 500),
                     sigma2 = c(0.1, 0.4, 0.5, 0.6, 0.7, 0.8, 1.0, 1.1, 1.2),
                     reps = 1000, seed = 20260911)
print(mc)
plot(mc, which = "identity")
```

The data-generating process is $RV_t = \mu_t\eta_t$ with $\eta_t$ gamma
with unit mean, so positivity is structural, the conditional mean is
exactly the HAR conditional mean, and no truncation or transformation is
required anywhere in the design.

A short run for illustration:

```{r}
mc <- tpr_montecarlo(n = 200, sigma2 = c(0.5, 1.0), reps = 10,
                     seed = 1, progress = FALSE)
print(mc)
```

The header of that output is the point of the exercise. The share of
replications in which `q` exceeded one must be 100 per cent and the
largest Proposition 2 residual must be of order `1e-16`, because both
are identities rather than empirical regularities.

## References

Corsi, F. (2009). A simple approximate long-memory model of realized
volatility. *Journal of Financial Econometrics* 7, 174--196.

Hoerl, A. E. and Kennard, R. W. (1970). Ridge regression: biased
estimation for nonorthogonal problems. *Technometrics* 12, 55--67.

Hoerl, A. E., Kennard, R. W. and Baldwin, K. F. (1975). Ridge
regression: some simulations. *Communications in Statistics* 4,
105--123.

Lipovetsky, S. and Conklin, W. M. (2005). Ridge regression in
two-parameter solution. *Applied Stochastic Models in Business and
Industry* 21, 525--540.
