---
title: "Two-parameter ridge for any regression"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Two-parameter ridge for any regression}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7,
                      fig.height = 4.5)
library(tpridge)
```

The package is built around volatility models, but the estimator is not.
Two-parameter ridge regression is a remedy for an ill-conditioned
design, and the design can come from anywhere. This vignette uses
`tpr_lm()`, the formula interface, on the data with which Hoerl and
Kennard illustrated ridge regression in the first place.

## The Longley data

Longley (1967) assembled sixteen annual observations on United States
employment against six macroeconomic series, as a test of the numerical
accuracy of regression software. The regressors are so nearly collinear
that early programs returned different answers from one another.

```{r}
data(longley)
m <- tpr_lm(Employed ~ GNP + Unemployed + Armed.Forces +
              Population + Year, data = longley)
m
```

```{r}
summary(m)
```

The diagnostics say what the problem is. The condition number of the
cross-product matrix is in the thousands, and several variance inflation
factors are far above the conventional threshold of ten. Gross national
product, population and the year index all measure much the same
expanding economy.

## What the second parameter does here

Notice the recovery figure in that summary: it is small. The reason is
visible in the same output. The data-driven penalty is mild, so `q` is
barely above one, and the fit gain is proportional to `(q-1)^2`. The
correction has almost nothing to work with.

That is not a failure of the method but a property of it, and it is
worth seeing directly. Fix the shrinkage rather than letting the data
choose it:

```{r}
for (df in c(5, 4, 3, 2, 1)) {
  f <- tpr_lm(Employed ~ GNP + Unemployed + Armed.Forces +
                Population + Year, data = longley,
              rule = "df", target_df = df)
  cat(sprintf("df %4.1f   q %7.4f   recovery %5.1f%%   R2 ridge-I %.5f\n",
              df, f$q, 100 * f$recovery, f$r2_in[["Ridge-I"]]))
}
```

As the penalty tightens, `q` grows and the share of the sacrificed fit
that the second parameter restores grows with it. This is the content of
Proposition 2: the gain is exactly
$(q-1)^2\|X\hat\beta_I\|^2/y'y$, so it is negligible when `q` is close
to one and substantial when it is not.

```{r}
plot(m, which = "path")
```

Least squares is flat, because it does not depend on the penalty. As
shrinkage increases the one-parameter estimator falls away while the
two-parameter estimator tracks least squares closely.

## Comparing with least squares

```{r}
ols <- lm(Employed ~ GNP + Unemployed + Armed.Forces + Population + Year,
          data = longley)
round(cbind(lm = coef(ols), coef(m, "OLS"),
            `Ridge-I` = coef(m, "Ridge-I"),
            `Ridge-II` = coef(m, "Ridge-II")), 4)
```

The first two columns agree, as they must: at the level of the
coefficients, `tpr_lm()` with no penalty is least squares. The ridge
columns pull the collinear coefficients towards one another, which is
the point.

## Prediction

```{r}
newd <- longley[1:3, ]
predict(m, newd, estimator = "Ridge-II")
predict(m, newd, estimator = "OLS")
```

## A note on inference

`summary()` prints no standard errors, and this is deliberate. The
sampling distribution of a shrinkage estimator is not that of least
squares, so the usual standard errors, $t$ statistics and confidence
intervals do not carry over. Reporting them would be misleading.

If inference is needed, a bootstrap that resamples the whole estimation
procedure, penalty selection included, is the defensible route. For
least squares coefficients on the same data, `lm()` and a
heteroskedasticity and autocorrelation consistent covariance estimator
remain appropriate.

## When is this worth using

The diagnostics in `summary()` answer that. A condition number in the
hundreds or thousands, or variance inflation factors above ten, indicate
a design where least squares coefficients are unstable and shrinkage is
worth considering. Whether the *second* parameter adds much then depends
on `q`, which the summary also reports: close to one and the correction
is negligible, materially above one and it recovers most of the fit that
one-parameter ridge gives up.

## References

Hoerl, A. E. and Kennard, R. W. (1970). Ridge regression: biased
estimation for nonorthogonal problems. *Technometrics* 12, 55--67.

Lipovetsky, S. and Conklin, W. M. (2005). Ridge regression in
two-parameter solution. *Applied Stochastic Models in Business and
Industry* 21, 525--540.

Longley, J. W. (1967). An appraisal of least squares programs for the
electronic computer from the point of view of the user. *Journal of the
American Statistical Association* 62, 819--841.
